Inducing Domain Specific Languages for Bayesian Program Learning

### 2018-01-31 paper.tex
started on list functions

We initially equip our DSL learner with addition and multiplication,
along with the possibility of introducing real-valued parameters, which we write as $\mathcal{R}$.
We define the likelihood of an observation $x$ by assuming a Gaussian noise model for the input/output examples and integrate over the real-valued parameters, which we collectively write as $\vec{\mathcal{R}}$:
\begin{align*}
\log   \probability\left[\{(i_n,o_n)\}|p\right] = \log \int \mathrm{d}\vec{\mathcal{R}}\; P_{\vec{\mathcal{R}}}(\vec{\mathcal{R}})\prod_{n\leq N}\mathcal{N}(p(i_n,\vec{\mathcal{R}})|o_n)
%\text{\emph{BIC:}} &\approx 
\end{align}
where $\mathcal{N}(\cdot |\cdot )$ is the normal density  and $P_{\vec{\mathcal{R}}}(\cdot )$ is a prior over $\vec{\mathcal{R}}$. We approximate this marginal using the BIC~\cite{Bishop:2006:PRM:1162264}:
\begin{align*}
  \log \probability[x|p]  \approx& \sum_{n\leq N}\log \mathcal{N}(p(i_n,\vec{\mathcal{R}}^*)|o_n) - \frac{D\log N}{2}
\end{align}
where $\vec{\mathcal{R}}^*$ is an assignment to $\vec{\mathcal{R}}$ found by performing gradient ascent on the likelihood of the observations w.r.t. $\vec{\mathcal{R}}$.

@@

\subsection{List problems}

>>>>

We initially equip our DSL learner with addition and multiplication,
along with the possibility of introducing real-valued parameters, which we write as $\mathcal{R}$.
We define the likelihood of an observation $x$ by assuming a Gaussian noise model for the input/output examples and integrate over the real-valued parameters, which we collectively write as $\vec{\mathcal{R}}$:
\begin{align*}
\log   \probability\left[\{(i_n,o_n)\}|p\right] = \log \int \mathrm{d}\vec{\mathcal{R}}\; P_{\vec{\mathcal{R}}}(\vec{\mathcal{R}})\prod_{n\leq N}\mathcal{N}(p(i_n,\vec{\mathcal{R}})|o_n)
%\text{\emph{BIC:}} &\approx 
\end{align*}
where $\mathcal{N}(\cdot |\cdot )$ is the normal density  and $P_{\vec{\mathcal{R}}}(\cdot )$ is a prior over $\vec{\mathcal{R}}$. We approximate this marginal using the BIC~\cite{Bishop:2006:PRM:1162264}:
\begin{align*}
  \log \probability[x|p]  \approx& \sum_{n\leq N}\log \mathcal{N}(p(i_n,\vec{\mathcal{R}}^*)|o_n) - \frac{D\log N}{2}
\end{align*}
where $\vec{\mathcal{R}}^*$ is an assignment to $\vec{\mathcal{R}}$ found by performing gradient ascent on the likelihood of the observations w.r.t. $\vec{\mathcal{R}}$.

@@

\subsection{List functions}
Our list function domain consists of tasks which are solved by functions
that take as input an integer or a list of integers, and have as output
either a Boolean, an integer, or a list of integers. Examples of these
functions are in Table~\ref{listexamples}. For each function, we create a
task $x$ by generating 15 input/output examples used for testing whether a
program produces the correct output. Supplying many examples reduces
ambiguity in the task's function, ensuring solutions achieve the desired
concept.

\begin{table}
\centering
\begin{tabular}{| l | l | l |}
  \hline
  \emph{name} & \emph{input} & \emph{output} \\
  \hline
  add-3 & [1\, 2\, 3\, 4] & [4\, 5\, 6\, 7] \\
  append-4 & [7\, 0\, 2] & [7\, 0\, 2\, 4] \\
  len & [3\, 5\, 12\, 1] & 4 \\
  range & 3 & [1\, 2\, 3] \\
  has-2 & [4\, 5\, 7\, 4] & \texttt{false} \\
  has-4 & [4\, 5\, 7\, 4] & \texttt{true} \\
  repeat-2 & [7\, 0] & [7\, 0\, 7\, 0] \\
  drop-3 & [0\, 3\, 8\, 6\, 4] & [6\, 4] \\
  \hline
\end{tabular}
\caption{Examples from the domain of list functions.}
\label{listexamples}
\end{table}

We supply \system with the DSL outlined in Table~\ref{listdsl}.

\begin{table}
\centering
\begin{tabular}{| l | l |}
  \hline
  \emph{name} & \emph{type} \\
  \hline
    empty & \texttt{tlist($\alpha$)} \\
    singleton & \texttt{arrow($\alpha$, tlist($\alpha$))} \\
    range & \texttt{int $\rightarrow$ tlist(int)} \\
    concat & \texttt{tlist($\alpha$) $\rightarrow$ tlist($\alpha$) $\rightarrow$ tlist($\alpha$)} \\
    map & \texttt{($\alpha$ $\rightarrow$ $\beta$) $\rightarrow$ tlist($\alpha$) $\rightarrow$ tlist($\beta$)} \\
    reduce & \texttt{($\beta$ $\rightarrow$ $\alpha$ $\rightarrow$ $\beta$) $\rightarrow$ $\beta$ $\rightarrow$ tlist($\alpha$) $\rightarrow$ $\beta$} \\
    \hline
    true & \texttt{bool} \\
    not & \texttt{bool $\rightarrow$ bool} \\
    and & \texttt{bool $\rightarrow$ bool $\rightarrow$ bool} \\
    or & \texttt{bool $\rightarrow$ bool $\rightarrow$ bool} \\
    \hline
    0, $\ldots$, 9 & \texttt{int} \\
    + & \texttt{int $\rightarrow$ int $\rightarrow$ int} \\
    * & \texttt{int $\rightarrow$ int $\rightarrow$ int} \\
    negate & \texttt{int $\rightarrow$ int} \\
    mod & \texttt{int $\rightarrow$ int $\rightarrow$ int} \\
    eq? & \texttt{int $\rightarrow$ int $\rightarrow$ bool} \\
    gt? & \texttt{int $\rightarrow$ int $\rightarrow$ bool} \\
    is-prime & \texttt{int $\rightarrow$ bool} \\
    is-square & \texttt{int $\rightarrow$ bool} \\
    sort & \texttt{tlist(int) $\rightarrow$ tlist(int)} \\
    \hline
    sum & \texttt{tlist(int) $\rightarrow$ int} \\
    reverse & \texttt{tlist($\alpha$) $\rightarrow$ tlist($\alpha$)} \\
    all & \texttt{($\alpha$ $\rightarrow$ bool) $\rightarrow$ tlist($\alpha$) $\rightarrow$ bool} \\
    any & \texttt{($\alpha$ $\rightarrow$ bool) $\rightarrow$ tlist($\alpha$) $\rightarrow$ bool} \\
    index & \texttt{int $\rightarrow$ tlist($\alpha$) $\rightarrow$ $\alpha$} \\
    filter & \texttt{($\alpha$ $\rightarrow$ bool) $\rightarrow$ tlist($\alpha$) $\rightarrow$ tlist($\alpha$)} \\
    slice & \texttt{int $\rightarrow$ int $\rightarrow$ tlist($\alpha$) $\rightarrow$ tlist($\alpha$)} \\
  \hline
\end{tabular}
\caption{DSL for the domain of list function.}
\label{listdsl}
\end{table}

\begin{figure}
  \includegraphics[width=\columnwidth]{figures/list.eps}
\end{figure}

We found that using a less sophisticated but equally-capable DSL made common
patterns, such as summation, unlikely and unlearnable in a small enumeration
bound.
